---
title: "Walk Roll Map Visualize"
---

Visualizations for [Walk Roll Map](https://www.walkrollmap.org/). 

* First draft is for infographics. 
* Later selected graphs may be added to website visualizations page. 
    * [BikeMaps.org](https://bikemaps.org/vis/) provides an example to follow.

```{r setup, echo=FALSE, warning=FALSE, message = FALSE, echo = FALSE, include = FALSE}

library(ggplot2)
library(leaflet)
library(leaflet.extras)
library(leaflet.opacity)
library(sf)
library(stringr)
library(tidyverse)
library(tm)
library(wordcloud)

# create a data folder (in the same directory) from here: https://www.dropbox.com/sh/e3xhul8zq0neko8/AACXmTYpRPjHNe6GwA73j7eia?dl=0
# run download_data.Rmd to update walk roll map data

# epsg code for wgs 84
geographic_projection <- 4326

# load data
wrm <- read_csv("data/wrm.csv")

wrm_spatial <- st_as_sf(wrm,
                        coords = c("lon", "lat"), 
                        crs = geographic_projection)

# timezones
timezones <- st_read("data/timezones/combined-shapefile.shp")
offsets <- read_csv("data/timezones/offsets.csv")

# colours and order to match the map

colour_pallete <- c("#FFC945",  # hazard: yellow
                    "#4281CD",  # missing amenity: blue
                    "#F8485E")  # incident: red

wrm_levels <- c("hazard-concern",
                "amenity",
                "incident")

```

# Map markers and heatmap
The heatmap might not make as much sense as it does for BikeMaps, because the 
reports on Walk Roll Map are more dispersed. Note that we can add opacity in a 
later step (using Javascript) so it looks exactly like on BikeMaps. We can also 
style the markers to match Walk Roll Map (this is a rough draft).

To do: connect the graphs to the map extent to spatially subset the data.

```{r map, echo = FALSE, warning = FALSE, message = FALSE}
# centre on Victoria
lat_centre <- 48.4295352
lon_centre <- -123.3596742
default_zoom = 12

hazard_icons <- awesomeIcons(
  icon = 'fa-solid fa-circle',
  iconColor = 'darkblue',
  library = 'fa',
  markerColor = 'lightgray'
)

amenity_icons <- awesomeIcons(
  icon = 'fa-solid fa-circle',
  iconColor = 'darkblue',
  library = 'fa',
  markerColor = 'blue'
)

incident_icons <- awesomeIcons(
  icon = 'fa-solid fa-circle',
  iconColor = 'darkblue',
  library = 'fa',
  markerColor = 'red'
)


wrm_spatial$descriptions <- paste0("<b>",stringr::str_to_title(wrm_spatial$type),": </b>", 
                       str_to_sentence(wrm_spatial$feature_type), "<br>",
                       "<b>Date: </b>", wrm_spatial$date, "<br>",
                       "<b>Id: </b>", wrm_spatial$id, "<br>",
                       "<b>Description: </b>", wrm_spatial$description)

marker_map <- leaflet() %>% 
  setView(lng = lon_centre, 
          lat = lat_centre, 
          zoom = default_zoom) %>%
  addTiles() %>%
  addAwesomeMarkers(
    data = wrm_spatial %>% filter(type %in% 'hazard-concern'),
    label = ~as.character(str_to_sentence(type)),
    popup = ~as.character(descriptions),
    icon = hazard_icons,
    clusterOptions = markerClusterOptions()) %>%
  addAwesomeMarkers(
    data = wrm_spatial %>% filter(type %in% 'amenity'),
    label = ~as.character(str_to_sentence(type)),
    popup = ~as.character(descriptions),
    icon = amenity_icons,
    clusterOptions = markerClusterOptions()) %>%
  addAwesomeMarkers(
    data = wrm_spatial %>% filter(type %in% 'incident'),
    label = ~as.character(str_to_sentence(type)),
    popup = ~as.character(descriptions),
    icon = incident_icons,
    clusterOptions = markerClusterOptions()) %>%
  addHeatmap(
    data = wrm_spatial,
    radius = 40,
    minOpacity = 0.3,
    group = "Heatmap") %>%
  addLayersControl(
    overlayGroups = c("Heatmap"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>%
  hideGroup("Heatmap")

marker_map
```

## Reports by type
Colours match Walk Roll Map legend.

```{r report_types, echo = FALSE, warning = FALSE, message = FALSE}
ggplot(wrm, aes(x=factor(type,
                         levels = wrm_levels), 
                fill=factor(type,
                         levels = wrm_levels))) +  
  geom_bar( ) +
  scale_fill_manual(values = colour_pallete ) +
  theme_minimal() +
  theme(legend.position="none",
        axis.title.x = element_blank(),
        axis.title.y = element_blank()) + 
  scale_x_discrete(breaks = c("hazard-concern",
                              "amenity",
                              "incident"),
                   labels = c("Hazard/Concern",
                              "Amenity",
                              "Incident"))


```

## Reports by day of week
```{r day_of_week, echo = FALSE, warning = FALSE, message = FALSE}

wrm$day <- format(wrm$date, "%a") %>%
  factor(levels = c("Mon",
                    "Tue",
                    "Wed",
                    "Thu",
                    "Fri",
                    "Sat",
                    "Sun"))

ggplot(wrm, aes(fill=factor(type,
                         levels = wrm_levels), 
                x=day)) + 
    geom_bar(position="stack", stat="count") + 
  scale_fill_manual(values = colour_pallete ) +
  theme_minimal() +
  theme(legend.position="none",
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
```


## Reports by hour
Need to consider timezones. Add UTC offset to get local timezone.

* timezones: https://github.com/evansiroky/timezone-boundary-builder/releases
* timezone offsets: http://joda-time.sourceforge.net/timezones.html
```{r hourly, echo = FALSE, warning = FALSE, message = FALSE}
timezones_with_offsets <- timezones %>%
  left_join(offsets,
            by = c("tzid" = "timezone"))

wrm_spatial_tz <- st_intersection(wrm_spatial,
                                    timezones_with_offsets)

wrm_spatial_tz$hour <- (as.numeric(format(wrm_spatial_tz$date, "%H")) + 
                          as.numeric(wrm_spatial_tz$offset_in_hours)) 

hour_lt_0 <- which(wrm_spatial_tz$hour < 0)
hour_gt_24 <- which(wrm_spatial_tz$hour > 24)

wrm_spatial_tz[hour_lt_0, ]$hour <- 24 + wrm_spatial_tz[hour_lt_0, ]$hour
wrm_spatial_tz[hour_gt_24, ]$hour <- wrm_spatial_tz[hour_gt_24, ]$hour - 24


wrm_spatial_tz$hour <- wrm_spatial_tz$hour %>%
  factor(levels = c("0",
                    "1",
                    "2",
                    "3",
                    "4",
                    "5",
                    "6",
                    "7",
                    "8",
                    "9",
                    "10",
                    "11",
                    "12",
                    "13",
                    "14",
                    "15",
                    "16",
                    "17",
                    "18",
                    "19",
                    "20",
                    "21",
                    "22",
                    "23",
                    "24"))

ggplot(wrm_spatial_tz, aes(fill=factor(type,
                         levels = wrm_levels), 
                x=hour)) + 
    geom_bar(position="stack", stat="count") + 
  scale_fill_manual(values = colour_pallete ) +
  theme_minimal() +
  theme(legend.position="none",
        axis.title.x = element_blank(),
        axis.title.y = element_blank())
```


## Reports by day of year

```{r day_of_year, echo = FALSE, warning = FALSE, message = FALSE}
wrm$doy <- format(wrm$date, "%j") %>%
  factor(levels = as.character(0:365 %>% str_pad(width = 3,
                                                 pad = 0)))
# count and add zeros
doy <- wrm %>%
  group_by(doy) %>%
  summarise(count = n())

doy <- table(wrm$doy)

doy <- data.frame(doy)
names(doy) <- c("DOY", "Count")

ggplot(doy, aes(x = DOY, y = Count)) + 
    geom_bar(stat = "identity") + 
  theme_minimal() +
  theme(legend.position="none",
        axis.title.x = element_blank()) +
  scale_x_discrete(name = "date", 
                   breaks = format(c(strptime("01/01/2021", "%d/%m/%Y"),
                                     strptime("01/02/2021", "%d/%m/%Y"),
                                     strptime("01/03/2021", "%d/%m/%Y"),
                                     strptime("01/04/2021", "%d/%m/%Y"),
                                     strptime("01/05/2021", "%d/%m/%Y"),
                                     strptime("01/06/2021", "%d/%m/%Y"),
                                     strptime("01/07/2021", "%d/%m/%Y"),
                                     strptime("01/08/2021", "%d/%m/%Y"),
                                     strptime("01/09/2021", "%d/%m/%Y"),
                                     strptime("01/10/2021", "%d/%m/%Y"),
                                     strptime("01/11/2021", "%d/%m/%Y"),
                                     strptime("01/12/2021", "%d/%m/%Y")), "%j"), 
                   labels = c("Jan",
                              "Feb",
                              "Mar",
                              "Apr",
                              "May",
                              "Jun",
                              "Jul",
                              "Aug",
                              "Sep",
                              "Oct",
                              "Nov",
                              "Dec"))
```


### Top reports
```{r top_reports, echo = FALSE, warning = FALSE, message = FALSE}
am_types <- wrm %>%
  filter(type %in% "amenity") %>%
  group_by(feature_type) %>%
  summarise(Amenity = n())

haz_types <- wrm %>%
  filter(type %in% "hazard-concern") %>%
  group_by(feature_type, feature_subtype) %>%
  summarise(Hazard_concern = n())

inc_types <- wrm %>%
  filter(type %in% "incident") %>%
  group_by(feature_type, feature_subtype) %>%
  summarise(Incidents = n())

am_map <- am_types %>%
  top_n(5) %>%
  rename(Report = feature_type,
         Count = Amenity) %>%
  mutate(Report = reorder(Report, Count)) %>%
  arrange(-Count) %>%
  mutate(type = "Amenity")


haz_map <- haz_types %>%
  group_by(feature_type) %>%
  summarise(Hazard_concern = sum(Hazard_concern)) %>%
  top_n(5) %>%
  rename(Report = feature_type,
         Count = Hazard_concern) %>%
  mutate(Report = reorder(Report, Count)) %>%
  mutate(type = "Hazard")

inc_map <- inc_types %>%
  group_by(feature_subtype) %>%
  summarise(Incidents = sum(Incidents)) %>%
  top_n(5) %>%
  rename(Report = feature_subtype,
         Count = Incidents) %>%
  mutate(Report = reorder(Report, Count)) %>%
  arrange(-Count) %>%
  mutate(type = "Incident")

top_5_all <- rbind(am_map,
                   haz_map,
                   inc_map)

top_5_all %>%
  ggplot(aes(x = Count, 
             y = Report,
             fill = as.factor(Report))) +
  geom_col() +
  #scale_fill_brewer(palette = "Set1") + 
  labs(y = NULL,
       x = NULL) +
  theme_minimal() +
  theme(legend.position="none") + 
  facet_wrap(~type, ncol = 1, scales = "free")

```
```{r amenity_word_freq, echo = FALSE, warning = FALSE, message = FALSE, include = FALSE}
am_map %>%
  ggplot(aes(Count, Report,
             fill = as.factor(Report))) +
  geom_col() +
  scale_fill_brewer(palette = "Set1") + 
  labs(y = NULL,
       x = NULL) +
  theme_minimal() +
  theme(legend.position="none")

```

```{r hazard_word_freq, echo = FALSE, warning = FALSE, message = FALSE, include = FALSE}
haz_map %>%
  ggplot(aes(Count, Report,
             fill = as.factor(Report))) +
  geom_col() +
  scale_fill_brewer(palette = "Set1") + 
  labs(y = NULL,
       x = NULL) +
  theme_minimal() +
  theme(legend.position="none")

```

```{r incident_word_freq, echo = FALSE, warning = FALSE, message = FALSE, include = FALSE}
inc_map %>%
  ggplot(aes(Count, Report,
             fill = as.factor(Report))) +
  geom_col() +
  scale_fill_brewer(palette = "Set1") + 
  labs(y = NULL,
       x = NULL) +
  theme_minimal() +
  theme(legend.position="none")

```


## Tree maps

Tree maps show proportions as a set of nested rectangles. The area is
proportional to the number of reports. 

There is a lot of information - maybe too much!

```{r treemap_overall, echo = FALSE, warning = FALSE, message = FALSE}
library(treemap)

wrm_types <- wrm %>%
  group_by(type, feature_type, feature_subtype) %>%
  summarise(All_Reports = n())

p <- treemap(wrm_types,
            index=c("type","feature_type", "feature_subtype"),
            vSize="All_Reports",
            type="index",
            palette = c(colour_pallete[2],
                        colour_pallete[1],
                        colour_pallete[3]),
            bg.labels=100,
            align.labels=list(
              c("center", "center"), 
              c("right", "bottom")
            )  
          )            
```

This is an interactive version of the tree map above:
```{r treemap_overall_interactive, echo = FALSE, warning = FALSE, message = FALSE}
library(d3treeR)

inter <- d3tree2(p,  rootname = "All reports" )
inter
```

### Amenity type tree map
```{r tree_map_amenity, echo = FALSE, warning = FALSE, message = FALSE}
ammenity <- treemap(am_types,
            index=c("feature_type"),
            vSize="Amenity",
            type="index",
            palette = "Set2",
            bg.labels=100,
            align.labels=list(
              c("center", "center"), 
              c("right", "bottom")
            )  
          )            
 
```

### Hazard type tree map
```{r tree_map_hazard, echo = FALSE, warning = FALSE, message = FALSE}
haz_tree <- treemap(haz_types,
            index=c("feature_type", "feature_subtype"),
            vSize="Hazard_concern",
            type="index",
            palette = "Set2",
            bg.labels=100,
            align.labels=list(
              c("center", "center"), 
              c("right", "bottom")
            )  
          )            
 
```

### Incident type tree map
```{r tree_map_incident, echo = FALSE, warning = FALSE, message = FALSE}
inc_tree <- treemap(inc_types,
            index=c("feature_type", "feature_subtype"),
            vSize="Incidents",
            type="index",
            palette = "Set2",
            bg.labels=100,
            align.labels=list(
              c("center", "center"), 
              c("right", "bottom")
            )  
          )            
```

## Frequency of keywords in description
This section analyses the free text description field.

* The reference for this section is the [tidytext book](https://www.tidytextmining.com/tidytext.html).
* The figures are directly out of the book.

### Most frequent words in desciption 
```{r nlp_overal, echo = FALSE, warning = FALSE, message = FALSE}
library(tidytext)
library(wordcloud2)
data(stop_words)

amenity_text <- wrm %>%
  filter(type %in% "amenity") %>%
  select(description) %>%
  unnest_tokens(word, description) %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE) 

amenity_top <- amenity_text %>%
  arrange(-n) %>%
  top_n(8) %>%
  mutate(word = paste0("amenity. ", word)) %>% # this to get the order due to repeated words. will replace
  mutate(word = reorder(word, n),
         type = "Amenity")

hazard_text <- wrm %>%
  filter(type %in% "hazard-concern") %>%
  select(description) %>%
  unnest_tokens(word, description) %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE) 

hazard_top <- hazard_text %>%
  arrange(-n) %>%
  top_n(8) %>%
  mutate(word = paste0("hazard. ", word)) %>% # this to get the order due to repeated words. will replace
  mutate(word = reorder(word, n),
         type = "Hazard")

incident_text <- wrm %>%
  filter(type %in% "incident") %>%
  select(description) %>%
  unnest_tokens(word, description) %>%
  anti_join(stop_words) %>%
  count(word, sort = TRUE) 

incident_top <- incident_text %>%
  arrange(-n) %>%
  top_n(8) %>%
  mutate(word = paste0("incident. ", word)) %>% # this to get the order due to repeated words. will replace
  mutate(word = reorder(word, n),
         type = "Incident")

top_words <- rbind(amenity_top, hazard_top, incident_top)
top_words <- top_words %>%
  group_by(type) %>%
  mutate(word = reorder(word, n))

words <- top_words$word %>% as.character()
clean_labels <- words %>%
  str_remove("amenity. ") %>%
  str_remove("hazard. ") %>%
  str_remove("incident. ")
  

top_words %>%
  ggplot(aes(x = word, 
             y = n,
             fill = as.factor(word))) +
  geom_col() +
  coord_flip() +
  #scale_fill_brewer(palette = "Set1") + 
  labs(y = NULL,
       x = NULL) +
  scale_x_discrete(breaks = words,
                   labels = clean_labels) + # this is to get the order right
  theme_minimal() +
  theme(legend.position="none") + 
  facet_wrap(~type, ncol = 1, scales = "free")

```

### Amenity
```{r wordcloud_amenity, echo = FALSE, warning = FALSE, message = FALSE}
#wordcloud2(amenity_text %>% top_n(100))
# the older version of wordcloud seems more reliable for R notebooks...
library(wordcloud)
library(tm)
library(RColorBrewer)
pal <- brewer.pal(6,"GnBu")

words <- amenity_text %>% top_n(100)

wordcloud::wordcloud(words = words$word,
                     freq = words$n,
                     colors = pal)
```

### Hazard/Concern
```{r wordcloud_hazard, echo = FALSE, warning = FALSE, message = FALSE}
words <- hazard_text %>% top_n(100)
pal <- brewer.pal(6,"YlGn")

wordcloud::wordcloud(words = words$word,
                     freq = words$n,
                     colors = pal)

```

### Incident
```{r wordcloud_incident, echo = FALSE, warning = FALSE, message = FALSE}
words <- incident_text %>% top_n(100)
pal <- brewer.pal(6,"OrRd")

wordcloud::wordcloud(words = words$word,
                     freq = words$n,
                     colors = pal)
```

### Term frequency and inverse document frequency
These graphs show which words are unique to each type of report.

```{r nlp_tf_idf, echo = FALSE, warning = FALSE, message = FALSE}
report_words <- wrm %>%
  unnest_tokens(word, description) %>%
  count(type, word, sort = TRUE) %>%
  anti_join(stop_words)

plot_reports <- report_words %>%
  bind_tf_idf(word, type, n) %>%
  mutate(type = factor(type, levels = wrm_levels))

plot_reports %>% 
  group_by(type) %>% 
  slice_max(tf_idf, n = 7) %>% 
  #ungroup() %>%
  mutate(word = reorder(word, tf_idf)) %>%
  ggplot(aes(tf_idf, word, fill = type)) +
  scale_fill_manual(values = colour_pallete ) +
  geom_col(show.legend = FALSE) +
  labs(x = "tf-idf", y = NULL) +
  facet_wrap(~type, ncol = 2, scales = "free") +
  theme_minimal()
```

